---
title: "Data_manipulation1"
author: "Evelyn Gakinya"
date: "9/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,warning=FALSE,
                      fig.width = 8, fig.height = 6)
```
```{r}
library(readr)
df<-read.csv("C:\\Users\\EVAH\\Desktop\\RSessions\\wafanyikazi.csv",header = TRUE, sep = ",")
df
```
```{r}
library(tidyverse)
```
##generating a tabular table for average income grouped by gender
```{r}
tab <- df %>% 
      group_by(Gender) %>% 
      summarise(Average_income = mean(Income, na.rm = T))
tab
```
##generating a tabular table for minimum age,maximum age and average income grouped by Gender
```{r}
tab1 <- df %>% 
        group_by(Gender) %>% 
        summarise(min_age = min(Age), max_age = max(Age), average_income = mean(Income))
tab1
```
###Grouping using 2 variables
##generating a tabular table for minimum age,maximum age and average income grouped by Gender ##and Department.
```{r}
tab2 <- df %>% 
        group_by(Gender,Department) %>% 
        summarise(min_age = min(Age), max_age = max(Age), average_income = mean(Income))
tab2
```

```{r}
tab5 <- df %>% 
        group_by(Marital_Status) %>% 
        summarise(mean.Income = mean(Income))
```

##generating a new variable ,converting the variable into percentage then rounding off to 0dp
```{r}
tab3 <- df %>% 
        group_by(Role) %>% 
        summarise(no.of.emp = n()) %>% 
        mutate(percent = (no.of.emp/sum(no.of.emp)) * 100) %>% 
        mutate(percent = round((no.of.emp/sum(no.of.emp)) * 100, 0))
tab3
```
##generating a new tabular table with increasing gender and  count
```{r}
tab4 <- df %>% 
        group_by(Gender, Role) %>% 
        summarise(count = n()) %>% 
        arrange(Gender,count)
tab4
```
##rename()
##rename(new name = old name)
##when renaming several variables: rename(new1 = old1,new2 = old2)
```{r}
new_varnames <- df %>% 
                rename(Umri = Age, Sex = Gender, Dept = Department)
new_varnames
```
##bind_cols() needs respondents to match across the different datasets.
#inner_join()
#right_join()
#left_join()
# Generate two datasets, 
#1. df1: All variables containing a letter 'e' and Sid

```{r}
df1 <- df %>% 
       select(Sid, contains("e"))
df1
```
#2. df2: Contains the rest of the variables including Sid 
```{r}
df2 <- df %>% 
       select(-contains("e"))
df2
```
##merging df1 and df2 using inner_join
```{r}
merged_df1 <- inner_join(df1, df2, by = "Sid")
merged_df1
```
##merging df1 and df2 using left_join
```{r}
merged_df2 <- left_join(df1, df2, by = "Sid")
merged_df2
```
``
##merging df1 and df2 using right_join
```{r}
merged_df3 <- right_join(df1,df2, by = "Sid")
merged_df3
```
#### Generate a dataset of females
```{r}
df_females <- df %>% 
              filter(Gender == "Female")
df_females
```
# Generate a dataset that merges females data and the original data
#inner_join
```{r}
data <- inner_join(df_females, df)
data
```
#leftjoin

```{r}
data1 <- left_join(df_females, df)
data1
```
#right_join
```{r}
data2 <- right_join(df_females,df)
data2
```
##full_join
```{r}
data3 <- full_join(df_females, df)
data3
```

